Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
refs

Whereas biomass can be remotely sensed and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
Although these stocks can be important--exceeding 100 t C ha^-1^ in some stands (Figs. 7d-e, S23-S25), this study is the first to synthesize deadwood data on a global scale [but see @cook-patton_mapping_2020 for young forests].
Unfortunately, data remain too sparse for statistical comparison across biomes (Figs. 7, S23-S25; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_2015]. 
Further research on non-living C stocks in the world's forests will be essential to completing the picture.

@@

Our study reveals that most C fluxes quickly increase to a plateau as stands age (Fig. 6), consistent with current understanding of age trends in forest C cycling [e.g., @anderson-teixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_auto$) would be minimal immediately following a stand-clearing disturbance.
These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. 7).
In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2020; @bond-lamberty_contribution_2004].
In this study, we detect no significant age trends in either variable.

@@

It is also interesting to consider the complementary utility of global-scale but spatially discontinuous databases such as ForC and remote wall-to-wall remote sensing products. The latter provide unparalleled insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bond-lamberty_estimating_2016; @anav_spatiotemporal_2016]. 
Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems [@konings_global_2019; @liu_detecting_2018].
Biomass is the largest C stock in most forests, and most of the emphasis has traditionally been on this variable. 
Remote-sensing driven biomass estimates [e.g., @saatchi_benchmark_2011], calibrated based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], are well suited for this task. 
Note, however, that factors such as stand age and disturbance history are difficult, if possible, to detect remotely, and can only be characterized for very recent decades [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018]. 
Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential [@luyssaert_old-growth_2008].

In contrast, carbon allocation within forest ecosystems and respiration fluxes cannot be remotely sensed. 
Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detecting_2018] typically compute respiration as residuals of all other terms [@bond-lamberty_estimating_2016; @harmon_heterotrophic_2011]. 
This means that the errors on respiration outputs are likely to be large and certainly poorly constrained, offering a unique opportunity for databases such as ForC and SRDB [@jian_restructured_2020] to provide observational benchmarks. 
For example, @konings_global_2019 produced a unique top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates [@bond-lamberty_new_2018] and used as an internal consistency check on other parts of the carbon cycle [@phillips_value_2016].

### Conclusions
As climate change accelerates, understanding and managing the carbon dynamics of forests--notably including dynamics and fluxes that cannot be observed by satellites--is critical to forecasting, mitigation, and adaptation.
The C data in *ForC*, as summarized here, will be valuable to these efforts.
Notably, the fact that tropical forests tend to have both the highest rates of C sequestration in young stands [Fig. 7; @cook-patton_mapping_2020], fueled by their generally high C flux rates (Table 1; Fig. 6), and the highest mean biomass [Fig. 7; Table 1; @hu_mapping_2016; @jian_reconstructed_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation [@grassi_key_2017], along with high-biomass old-growth temperate stands [@goldstein_protecting_2020].
It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$; @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that could not be recovered on a time scale relevant to climate change mitigation [@goldstein_protecting_2020].
Generally speaking, the conservation of mature forests will yield greater climate benefits [@anderson-teixeira_greenhouse_2011], but both approaches are critical to avoiding catastrophic climate change [@ipcc_global_2018].

>>>>

Whereas biomass can be remotely sensed and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
Although these stocks can be important--exceeding 100 t C ha^-1^ in some stands (Figs. 7d-e, S23-S25), this study is the first to synthesize deadwood data on a global scale [but see @cook-patton_mapping_2020 for young forests].
Unfortunately, data remain too sparse for statistical comparison across biomes (Figs. 7, S23-S25; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_global_2005]. 
Further research on non-living C stocks in the world's forests will be essential to completing the picture.

@@

Our study reveals that most C fluxes quickly increase to a plateau as stands age (Fig. 6), consistent with current understanding of age trends in forest C cycling [e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_auto$) would be minimal immediately following a stand-clearing disturbance.
These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. 7).
In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2020; @bond-lamberty_contribution_2004].
In this study, we detect no significant age trends in either variable.

@@

It is also interesting to consider the complementary utility of global-scale but spatially discontinuous databases such as ForC and remote wall-to-wall remote sensing products. The latter provide unparalleled insight into aboveground carbon stocks, but less constraint on belowground stocks or carbon fluxes in general [@bondlamberty_estimating_2016; @anav_spatiotemporal_2015]. 
Combining observational data and remote observations may provide a much more comprehensive and accurate picture of global forest C cycling, particularly when used in formal data assimilation systems [@konings_global_2019; @liu_detecting_2018].
Biomass is the largest C stock in most forests, and most of the emphasis has traditionally been on this variable. 
Remote-sensing driven biomass estimates [e.g., @saatchi_benchmark_2011], calibrated based on high-quality ground-based data [@schepaschenko_forest_2019; @chave_ground_2019], are well suited for this task. 
Note, however, that factors such as stand age and disturbance history are difficult, if possible, to detect remotely, and can only be characterized for very recent decades [@hansen_high-resolution_2013; @song_global_2018; @curtis_classifying_2018]. 
Ground-based data such as *ForC* are therefore valuable in defining age-based trajectories in biomass, as in @cook-patton_mapping_2020, and thus constraining variables such as carbon sink potential [@luyssaert_old-growth_2008].

In contrast, carbon allocation within forest ecosystems and respiration fluxes cannot be remotely sensed. 
Efforts such as the Global Carbon Project [@friedlingstein_global_2019] and NASA’s Carbon Monitoring System [@liu_detecting_2018] typically compute respiration as residuals of all other terms [@bondlamberty_estimating_2016; @harmon_heterotrophic_2011]. 
This means that the errors on respiration outputs are likely to be large and certainly poorly constrained, offering a unique opportunity for databases such as ForC and SRDB [@jian_restructured_2020] to provide observational benchmarks. 
For example, @konings_global_2019 produced a unique top-down estimate of global heterotrophic respiration that can both be compared with extant bottom-up estimates [@bondlamberty_new_2018] and used as an internal consistency check on other parts of the carbon cycle [@phillips_value_2017].

### Conclusions
As climate change accelerates, understanding and managing the carbon dynamics of forests--notably including dynamics and fluxes that cannot be observed by satellites--is critical to forecasting, mitigation, and adaptation.
The C data in *ForC*, as summarized here, will be valuable to these efforts.
Notably, the fact that tropical forests tend to have both the highest rates of C sequestration in young stands [Fig. 7; @cook-patton_mapping_2020], fueled by their generally high C flux rates (Table 1; Fig. 6), and the highest mean biomass [Fig. 7; Table 1; @hu_mapping_2016; @jian_restructured_2020] reinforces the concept that conservation and restoration of these forests is a priority for climate change mitigation [@grassi_key_2017], along with high-biomass old-growth temperate stands [@goldstein_protecting_2020].
It is also important to note the trade-off in climate mitigation potential of restoration of young forests, with high rates of CO~2~ sequestration [$NEP$; @cook-patton_mapping_2020], versus conservation and management of mature forests, with low $NEP$ but high C stocks that could not be recovered on a time scale relevant to climate change mitigation [@goldstein_protecting_2020].
Generally speaking, the conservation of mature forests will yield greater climate benefits [@anderson-teixeira_greenhouse_2011], but both approaches are critical to avoiding catastrophic climate change [@ipcc_global_2018].
